Graph Topological Aspects of Granger Causal Network Learning

### 2019-07-12 paper/paper.tex
Remove todo note.
Alphabetized keywords.
Amend keywords.
Breaking paper into separate pieces of work.  Also fixed significant error in one of the Lemmas.

% TODO: read this https://lemire.me/blog/rules-to-write-a-good-research-paper/

@@

\keywords{Granger-causality, causality graph, time series, vector autoregression,
  LASSO, gene regulatory networks, network learning, EEG}

@@

\subsection{Contributions}
In the context of time series data, sparsity assumptions remain
important, but there is significant additional structure that may
arise as a result of considering the topology of the underlying
Granger-causality graph, which to our knowledge remains largely
unexplored.  The focus of this paper is to shed light on some of these
topological questions, in particular, we study a particularly simple
notion of causality graph topology which we term ``strongly causal''
and show that $\VAR$ models having this structure satisfy natural
intuitive notions of ``information flow'' through the graph.
Moreover, we show that such graphs are perfectly recoverable with only
\textit{pairwise} Granger-causality tests, which would otherwise suffer
from serious confounding problems.  Our finite sample results are
based on simulations, where we show promising results for these graph
topologies where our algorithm performs substantially better in our
simulation setup than do competing LASSO algorithms, even for graphs
that do not exactly satisfy our strongly-causal topology assumptions.

In the case of gene expression networks, we show examples from the
literature which suggest our concept of a ``strongly causal graph''
topology may have application in this field (see Section
\ref{sec:strongly_causal_graphs})

>>>>

@@

\keywords{causality graph, EEG, gene regulatory networks,
  Granger-causality, LASSO, network learning, time series, vector
  autoregression}

@@

\subsection{Contributions}
In the context of time series data, sparsity assumptions remain
important, but there is significant additional structure that may
arise as a result of considering the topology of the underlying
Granger-causality graph, which to our knowledge remains largely
unexplored.  The focus of this paper is to shed light on some of these
topological questions, in particular, we study a particularly simple
notion of causality graph topology which we term ``strongly causal''
and show that stationary times series whose underlying causality graph
has this structure satisfy natural intuitive notions of ``information
flow'' through the graph.  Moreover, we show that such graphs are
perfectly recoverable with only \textit{pairwise} Granger-causality
tests, which would otherwise suffer from serious confounding problems.
Our finite sample results are based on simulations, where we show
promising results for these graph topologies where our algorithm
performs substantially better in our simulation setup than do
competing LASSO algorithms, even for graphs that do not exactly
satisfy our strongly-causal topology assumptions.

In the case of gene expression networks, we show examples from the
literature which suggest our concept of a ``strongly causal graph''
topology may have application in this field (see Section
\ref{sec:strongly_causal_graphs}).
